Parity games in pushdown systems with pebbles

### 2023-06-11 reachability/article.tex
up

Our main result states that
deciding
the winner of a parity game on a graph of a pebble pushdown system
is decidable but nonelementary.
The contribution is two-fold.

	For the nonelementary lower bound,
we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
from \cite{Sto74}, to the
problem of deciding whether a player has a winning strategy.

	For the main contribution, we use pebble alternating two-way tree automata to solve pebble pushdown parity games.
Alternating two-way tree automata with parity winning conditions have been used in the past to solve parity games on pushdown graphs \cite{cachat2002two}. We show that pebble alternating two-way tree automata can similarily be used to solve parity games on pebble pushdown graphs.
We then prove  decidability 
by 
a reduction
 to the problem of model-checking $\MSO$ on a well chosen structure, whose decidability is
 guaranteed by
Muchnik's Theorem~\cite{Wal96}.

>>>>

Our main result consists in
proving parametric pushdown reachability games
and parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
$n$ is fixed,
and providing a nonelementary lower bound for parametric pushdown reachability games in general.

	For the nonelementary lower bound,
we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
from \cite{Sto74}, to the
problem of deciding whether a player has a winning strategy for parametric pushdown reachability games.

	For the upper bound, we start by replacing parameters by pebbles acting as registers, leading to a more general model, pebble pushdown automata.
 We then show that higher-order pushdown automata parity games can be used to solve parity games on
 the transition systems of pebble pushdown automata,
using one additional stack level for each pebble.
 Since solving parity games on higher-order pushdown automata with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown automata with $n$ parameters.
